---
title: "Crime dataset"
---
## Overall description

The crime data was collected by FBI's Uniform Crime Reporting (UCR) Program, which recorded different types of offenses by states, regions, and geographic divisions. All offenses were classified to two main categories:

* **Violent crime**

The violent crime involved force or threat of force against the victims. Here it was divided into four subtypes of offenses:

 * Murder and nonnegligent manslaughter
 * Rape
 * Robbery
 * Aggravated assault

* **Property crime**

The property crime indicated offenses with property lost and without force or threat of force. Here it included three subtypes:

 * Burglary
 * Larceny-theft
 * Motor vehicle theft


```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(PerformanceAnalytics)
library(rvest)
library(httr)
library(gridExtra)
library(flexdashboard)
library(viridis)
library(plotly)
library(gsheet)
```



```{r, include = FALSE}
options( warn = -1 )
# load data and clean up column names
# not to omit NAs
crime_raw = gsheet2tbl('https://docs.google.com/spreadsheets/d/1iYhSFxs_IdRDrfrn1xqdK6uesNKWRjCIJ2GYloO0Ll8/edit#gid=249460096') 
crime_raw = crime_raw[3:201,]

crime = crime_raw %>% 
  janitor::clean_names() %>% 
    rename_(population = names(.)[3],
          violent_crime_number = names(.)[4],
          violent_crime_rate = names(.)[5],
          murder_manslaughter_number = names(.)[6],
          murder_manslaughter_rate = names(.)[7],
          rape_revise_number = names(.)[8],
          rape_revise_rate = names(.)[9],
          rape_legacy_number = names(.)[10],
          rape_legacy_rate = names(.)[11],
          robbery_number = names(.)[12],
          robbery_rate = names(.)[13],
          aggravated_assault_number = names(.)[14],
          aggravated_assault_rate = names(.)[15],
          property_crime_number = names(.)[16],
          property_crime_rate = names(.)[17],
          burglary_number = names(.)[18],
          burglary_rate = names(.)[19],
          larceny_theft_number = names(.)[20],
          larceny_theft_rate = names(.)[21],
          motor_vehicle_theft_number = names(.)[22],
          motor_vehicle_theft_rate = names(.)[23]) %>% 
  mutate(area = as.factor(area)) 

# Remove comma and convert to numeric
crime[,3:23] <- as.numeric(gsub(",","",unlist(crime[,3:23])))

crime[,2:23] <- as.numeric(unlist(crime[,2:23]))

# fill up area name
i = 1
for (i in 0:65) {
  n = 3 * i + 1
  crime$area[n + 1] = paste0(crime$area[n]) 
  crime$area[n + 2] = paste0(crime$area[n])
  i = i + 1
}

crime  <- crime %>% 
  na.omit()

crime$year = as.factor(crime$year)


index_ne <- which(crime$area == "Northeast")
idnex_mw <- which(crime$area == "Midwest")
idnex_s <- which(crime$area == "South")
idnex_w <- which(crime$area == "West")

index <- c(1:length(crime$area))
crime$region <- c(rep(0, length(crime$area)))

index_ne_state <- which(index > max(index_ne) & index < min(idnex_mw))
crime$region[index_ne_state] <- 1

index_ne_state <- which(index > max(idnex_mw) & index < min(idnex_s))
crime$region[index_ne_state] <- 2

index_ne_state <- which(index > max(idnex_s) & index < min(idnex_w))
crime$region[index_ne_state] <- 3

index_ne_state <- which(index > max(idnex_w) & index < 133)
crime$region[index_ne_state] <- 4

crime[,4:23] <- crime[,4:23]/100

```



```{r data cleaning on crime dataset, include = FALSE}
options( warn = -1 )
# Some further clean up on crime data 
# Filter out "percent change" in "year" variable
crime_clean = 
  crime %>%
  filter(year != "NA")

# Create a new dataframe to store the data of different large areas
crime_largearea = 
  crime_clean %>%
  filter(area %in% c("United States Total", "Northeast", "Midwest", "South", "West"))

# Create subtotal for each big areas
crime_northeast = 
  crime_clean %>%
  filter(area %in% c("New England", "Middle Atlantic"))

crime_midwest = 
  crime_clean %>%
  filter(area %in% c("East North Central", "West North Central"))

crime_south = 
  crime_clean %>%
  filter(area %in% c("South Atlantic", "East South Central", "West South Central"))

crime_west = 
  crime_clean %>%
  filter(area %in% c("Mountain", "Pacific", "Puerto Rico"))


# Filter out these areas in the tidied dataset
crime_clean = 
  crime_clean %>%
  filter(!area %in% c("United States Total", "Northeast", "Midwest", "South", "West", "New England", "Middle Atlantic", "East North Central", "West North Central", "South Atlantic", "East South Central", "West South Central", "Mountain", "Pacific", "Puerto Rico"))



```



## Violent Crime Rate 


```{r, echo=FALSE}
options( warn = -1 )
# Group by districts(i.e. Northeast, Midwest, South, West), each district can be subdivide into smaller categories
# Rank areas by average violent crime rate from 2015 to 2016
crime_plot <- crime %>% 
  filter(!area %in% c("United States Total", "Northeast", "Midwest", "South", "West", "New England", "Middle Atlantic", "East North Central", "West North Central", "South Atlantic", "East South Central", "West South Central", "Mountain", "Pacific", "Puerto Rico")) %>%  
  filter(year != "Percent change") %>%
  group_by(area) %>%
  summarise_all(funs(mean))

crime_plot$region <- as.character(crime_plot$region)
index_1 <- which(crime_plot$region == "1")
crime_plot$region[index_1] = "Northeast"
index_2 <- which(crime_plot$region == "2")
crime_plot$region[index_2] = "Midwest"
index_3 <- which(crime_plot$region == "3")
crime_plot$region[index_3] = "South"
index_4 <- which(crime_plot$region == "4")
crime_plot$region[index_4] = "West"
crime_plot$region <- as.factor(crime_plot$region)

crime_plot_violent_about <- crime_plot %>% 
  mutate(area = fct_reorder(area, violent_crime_rate)) %>% 
  ggplot(aes(x = area, y = violent_crime_rate, fill = region)) +
  geom_bar(stat = "identity", width = 0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 8),
        title = element_text(size = 14)) +
  labs(
    x = "State",
    y = "Violent crime rate (Per 1000)"
  )
crime_plot_violent_about
ggsave(filename = "violent_crime_rate_overall_plot.png", plot = crime_plot_violent_about, width = 30, height = 20, units = "cm")


```

The highest violent crime rate was observed in District of Columbia, with 12 victims per 1000 people in one year. This rate was 2 to 3 times the violent crime rate of most states, indicating the potential insecurity regarding violent offences in District of Columbia. Moreover, states in the south region tended to have higher violent crime rates compare to other regions.


## Property crime

```{r, echo = FALSE, warning = FALSE}
crime_plot_property_about <- crime_plot %>% 
  mutate(area = fct_reorder(area, property_crime_number)) %>% 
  ggplot(aes(x = area, y = property_crime_number, fill = region)) +
  geom_bar(stat = "identity", width = 0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 8),
        title = element_text(size = 14)) +
  labs(
    x = "State",
    y = "Property crime rate (Per 1000)"
  )
crime_plot_property_about
```

The highest property crime rates were seen in Florida, Texas, and California, which doubled or tripled the crime rate of other states. The possible reasons could be large population and huge influx to tourist draw during vacations.

On the other hand, District of Columbia, which was notorious for the high violent crime rate, had a very low property crime rate. Therefore, it could be reasonable to conclude that wealthier states tended to have higher property crime rates.




