---
title: "Drug Code"
---

## Overall description

The drug data was collected by The National Survey on Drug Use and Health (NSDUH), an annual survey conducted in all 50 states and the District of Columbia. There were approximately 70,000 participants and questions on tobacco, alcohol, and drug use, mental health and other health-related issues were asked. Our study focued on substance abuse (alcohol, marijuana, cocaine, and illicit drugs) in state level.



```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(PerformanceAnalytics)
library(rvest)
library(httr)
library(gridExtra)
library(flexdashboard)
library(viridis)
library(plotly)
library(gsheet)
```


```{r read_data, echo=FALSE, message=FALSE}
drug_raw =
  read_csv("https://think.cs.vt.edu/corgis/csv/drugs/drugs.csv?forcedownload=1/drugs.csv") %>%
  janitor::clean_names()
```


```{r clean_data,echo=FALSE,warning=FALSE,message=FALSE}
drug_data = 
  drug_raw %>%
  select(year, state,
         rates_alcohol_abuse_past_year_12_17:rates_alcohol_dependence_past_year_26,
         rates_alcohol_need_treatment_past_year_12_17:rates_illicit_drugs_abuse_past_month_26, 
         rates_illicit_drugs_cocaine_used_past_year_12_17:rates_illicit_drugs_need_treatment_past_year_26,
         rates_marijuana_new_users_12_17:rates_marijuana_used_past_year_26) %>%
  gather(key, value, -state, -year) %>% 
  mutate(key = str_replace_all(key, c("year_" = "year:", "month_" = "month:", "risk_" = "risk:", "users_" = "users:"))) %>%
  separate(key, into = c("rate", "age"), sep = ":", extra = "merge") %>%
  mutate(rate = str_replace(rate, "^rates_", ""))
```

## Trend of different drug uses by age groups


```{r line_plot, echo = FALSE, warning = FALSE}
# Make a line plot 
drug_data %>%     
filter(!str_detect(rate, "perceptions"),
         !str_detect(rate, "past_month"),
         !str_detect(rate, "new_user")) %>%
  mutate(rate = str_replace(rate, "alcohol_", "alcohol:"),
         rate = str_replace(rate, "illicit_drugs_", "illicit_drugs:"),
         rate = str_replace(rate, "marijuana_", "marijuana:")) %>% 
  separate(rate, into = c("drug", "other"), by = ":") %>% 
  select(-other) %>%
  group_by(year, drug, age) %>% 
  summarize(rate = sum(value)) %>% 
  ggplot(aes(x = year, y = rate, 
             color = drug, 
             group = interaction(age, drug))) +
  geom_line() +
  geom_point(aes(shape = age)) +
  theme_bw() +
  labs(
    title = "Different drug uses in 2002-2014 by age groups",
    x = "Year",
    y = "Rate of drug use (in thousands)",
    caption = "Data from the drug dataset"
  ) +
   theme(legend.position = "bottom") +
  scale_x_continuous(breaks = c(2002, 2004, 2006, 2008, 2010, 2012, 2014))
```

The highest rates of alcohol, illicit drugs, and marijuana were among people aged 18 to 25. There was a decreasing trend in alcohol abuse among people aged 12 to 25. The rates of both illicit drugs--cocaine and marijuana, in general, remained constant throughout years.



## Trend of perception risk in drug use by age groups

```{r barplot, echo=FALSE, warning=FALSE}
drug_data %>% 
  filter(str_detect(rate, "perception")) %>% 
  group_by(year, rate, age) %>% 
  summarize(mean_rate = mean(value)) %>% 
  ggplot(aes(x = year, y = mean_rate)) +
  geom_bar(aes(fill = age), stat = "identity", width = 0.6) +
  ylim(0, 350) +
  theme_bw() +
  labs(
    title = "Perception risk in 2002-2014 by age groups",
    x = "Year",
    y = "Rate for perception of risk",
    caption = "Data from the drug dataset"
  ) +
   theme(legend.position = "bottom") +
  scale_x_continuous(breaks = c(2002, 2004, 2006, 2008, 2010, 2012, 2014))
```

The overall perception of risk was decreasing as shown in the plot, especially after year of 2010, the rate reduced continously which indicated that more and more people did not realize the risk of drug abuse. Among the age groups, people who were 12-17 years old were more likely to ignore the harm of drug use. These results might result from lack of education in drug in school and the downtrend of perception of risk as well as lowering trend of drug users' age set an alarm for potential deterioration of drug use.


