---
title: "Explore possible relationship between crime rate and drug use"
output:
  html_document: 
    toc: true
    toc_float: true
---

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(PerformanceAnalytics)
library(rvest)
library(httr)
library(DT)
library(gsheet)
```


```{r, include=FALSE, warning=FALSE, message=FALSE}
#Read in Crime data
crime = readxl::read_excel("./data/crime-data-2016.xls", range = ("A4:W202")) %>% 
  janitor::clean_names() %>% 
    rename_(population = names(.)[3],
          violent_crime_number = names(.)[4],
          violent_crime_rate = names(.)[5],
          murder_manslaughter_number = names(.)[6],
          murder_manslaughter_rate = names(.)[7],
          rape_revise_number = names(.)[8],
          rape_revise_rate = names(.)[9],
          rape_legacy_number = names(.)[10],
          rape_legacy_rate = names(.)[11],
          robbery_number = names(.)[12],
          robbery_rate = names(.)[13],
          aggravated_assault_number = names(.)[14],
          aggravated_assault_rate = names(.)[15],
          property_crime_number = names(.)[16],
          property_crime_rate = names(.)[17],
          burglary_number = names(.)[18],
          burglary_rate = names(.)[19],
          larceny_theft_number = names(.)[20],
          larceny_theft_rate = names(.)[21],
          motor_vehicle_theft_number = names(.)[22],
          motor_vehicle_theft_rate = names(.)[23]) %>% 
  mutate(area = as.factor(area)) 
crime[,2:23] <- as.numeric(unlist(crime[,2:23]))

# fill up area name
i = 1
for (i in 0:65) {
  n = 3 * i + 1
  crime$area[n + 1] = paste0(crime$area[n]) 
  crime$area[n + 2] = paste0(crime$area[n])
  i = i + 1
}

crime  <- crime %>% 
  na.omit()

crime$year = as.factor(crime$year)

# Filter out "percent change" in "year" variable
crime_clean = 
  crime %>%
  filter(year != "NA")


# Change the name of "area" to "state"
colnames(crime_clean)[1] = "state"
```



```{r drug dataset load and clean up, include=FALSE, warning=FALSE, message=FALSE}
# Remove the block later
drug_raw =
  read_csv("https://think.cs.vt.edu/corgis/csv/drugs/drugs.csv?forcedownload=1/drugs.csv") %>%
  janitor::clean_names()

drug_clean = 
  drug_raw %>%
  select(year, state,
         rates_alcohol_abuse_past_year_12_17:rates_alcohol_dependence_past_year_26,
         rates_alcohol_need_treatment_past_year_12_17:rates_illicit_drugs_abuse_past_month_26, 
         rates_illicit_drugs_cocaine_used_past_year_12_17:rates_illicit_drugs_need_treatment_past_year_26,
         rates_marijuana_new_users_12_17:rates_marijuana_used_past_year_26) %>%
  gather(key, value, -state, -year) %>% 
  mutate(key = str_replace_all(key, c("year_" = "year:", "month_" = "month:", "risk_" = "risk:", "users_" = "users:"))) %>%
  separate(key, into = c("rate", "age"), sep = ":", extra = "merge") %>%
  mutate(rate = str_replace(rate, "^rates_", ""))

population_data = 
  drug_raw %>% 
  select(x12_17, x18_25:x26, year, state) %>% 
  gather(key = "age", value = "popul_in_thousands", -state, -year) %>% 
  mutate(age = str_replace(age, "x", ""))
```



## Variable Distribution Check

We first checked whether each columns (including both outcomes and predictors) were normally distribted.


 * Outcomes: rates of 10 types of crimes(two main crimes: violent and property, and their subtypes)
 
 * Predictors: use of 3 types of drugs, ratio of white race(possible confounders), and high school graduations(possible confounders)

```{r, include=FALSE, warning=FALSE, message=FALSE}
crime_mean_rate <- 
  crime_clean %>% 
  group_by(state) %>% 
  select(grep("rate",names(.))) %>% 
  summarise_all(funs(mean))
crime_mean_rate[,2:11] <- crime_mean_rate[,2:11]/100
```


```{r prepare the outcome and predictors, include=FALSE, warning=FALSE, message=FALSE}
# clean drug data
drug_mean_rate = drug_clean %>% 
  group_by(state, rate, age) %>% 
  summarize(drug_average = mean(value)) %>% 
  group_by(state, rate) %>% 
  summarize(drug_sum = sum(drug_average)) %>% 
  filter(!str_detect(rate, 'perceptions'),
         str_detect(rate, 'past_year')) %>% 
  mutate(rate = str_replace_all(rate, c("alcohol_" = "alcohol:", "illicit_drugs_" = "illicit_drugs:", "marijuana_" = "marijuana:"))) %>% 
  separate(rate, into = c("drug", "degree"), sep = ":", extra = "merge") %>% 
  select(-degree) %>% 
  group_by(state, drug) %>% 
  summarize(use_past_year = sum(drug_sum)) %>% 
# up to here, it's already a tidy dataset, but for regression we need to define each drug as a variable, so we choose to spread "drug" column
  spread(key = drug, value = use_past_year)

# read data for race proportion
race_data_raw = read_csv("./data/race_data.csv") %>% janitor::clean_names() %>%
  select(state, white)

# Combine X and Y into one data frame
regression_data_raw = left_join(crime_mean_rate, drug_mean_rate) %>% 
  select(-rape_legacy_rate)

regression_data_raw = left_join(regression_data_raw, race_data_raw)
```



```{r read in new dataset, include=FALSE, warning=FALSE, message=FALSE}
# Import dataset that contains education information
url = "https://en.wikipedia.org/wiki/List_of_U.S._states_by_educational_attainment"
education = read_html(url)

# tidy and clean
education_data = education %>% 
  html_nodes(css = "table") %>%
  .[[1]] %>% 
  html_table(fill = TRUE) %>% 
  janitor::clean_names() %>% 
  select(state, percent_high_school_graduate) %>% 
  filter(state != "United States") %>% 
  mutate(percent_high_school_graduate = str_replace(percent_high_school_graduate, "%", "")) %>% 
  mutate(percent_high_school_graduate = as.numeric(percent_high_school_graduate))

regression_data_all = left_join(regression_data_raw, education_data)

```


```{r distribution checking, echo = FALSE, warning=FALSE}
regression_data_all %>%
  select(-state) %>%
  chart.Correlation()
```

All columns were approximately normal distributed, which could be used for downstream analysis without transformation.

## Estimated Coefficients with Leverages

Fit multiple linear regression for all crime types and drug uses, as well as race and education. Check diagnostics plots to find possible outliers and influential points.

```{r linear regression, echo=FALSE, warning=FALSE, message=FALSE}
# MLR linear regression using list columns
significant_coefficients1 =
  regression_data_all %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  group_by(crime_type) %>%
  nest() %>%
  mutate(
    linear_model = map(data, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = .x)),
    linear_model = map(linear_model, broom::tidy)) %>%
  select(-data) %>%
  unnest() %>% 
  filter(p.value < 0.05,
         term != "(Intercept)")
  

# Make a new dataframe to store the result of regression
result_with =
  regression_data_all %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  group_by(crime_type) %>%
  nest() %>%
  mutate(
    linear_model = map(data, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = .x)),
    linear_model = map(linear_model, broom::tidy)) %>%
  select(-data) %>%
  unnest() %>%
  mutate_if(is.numeric, round, 2) 

datatable(result_with, options = list(pageLength = 5))
```

## Diagnostics for Models with Leverages

```{r model diagnostics, echo=FALSE, warning=FALSE, message=FALSE}
# Fit models and then check the assumption
model1 = lm(murder_manslaughter_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = regression_data_all)
summary(model1)
par(mfrow = c(2,2))
plot(model1)

model2 = lm(robbery_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = regression_data_all)
summary(model2)
par(mfrow = c(2,2))
plot(model2)
```

It was shown that both murder/manslaughter and robbery rate were positively related to marijuana use. In addition, murder/manslaughter rate was also related to illicit drugs. Therefore we would investigate deeper into MLR models for murder/manslaughter and robbery.

Two outliers/influential points were identified: District of Columbia(observation 9) and Hawaii(observation 14), the MLR model would be refitted without these two observations.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
murder_manslaughter_bt <- regression_data_all %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  filter(crime_type == "murder_manslaughter_rate") %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    linear_model = map(strap, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + percent_high_school_graduate, data = .x)),
    results = map(linear_model, broom::tidy)) %>% 
  select(results) %>%
  unnest()

robbery_rate_bt <- regression_data_all %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  filter(crime_type == "robbery_rate") %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    linear_model = map(strap, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + percent_high_school_graduate, data = .x)),
    results = map(linear_model, broom::tidy)) %>% 
  select(results) %>%
  unnest()
```


## Estimated Coefficients without Leverages

Then two outliers were removed and the model was fitted again

```{r linear regression_another, echo=FALSE, warning=FALSE, message=FALSE}
# MLR linear regression using list columns
regression_data_no2states <- regression_data_all %>% 
  filter(!state %in% c("District of Columbia", "Hawaii"))

significant_coefficients2 = 
  regression_data_no2states %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  group_by(crime_type) %>%
  nest() %>%
  mutate(
    linear_model = map(data, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = .x)),
    linear_model = map(linear_model, broom::tidy)) %>%
  select(-data) %>%
  unnest() %>% 
  filter(p.value < 0.05,
         term != "(Intercept)")

result_without = 
  regression_data_no2states %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  group_by(crime_type) %>%
  nest() %>%
  mutate(
    linear_model = map(data, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = .x)),
    linear_model = map(linear_model, broom::tidy)) %>%
  select(-data) %>%
  unnest() %>%
  mutate_if(is.numeric, round, 2)

datatable(result_with, options = list(pageLength = 5))
```

By removing these two observations, the estimated coefficients for drug use become not significant.


## Diagnostics for Models without Leverages

Check the diagnostics plots for MLR without two outliers/influential points.

```{r model diagnostics_another, echo=FALSE, warning=FALSE, message=FALSE}
# Fit models and then check the assumption
model1 = lm(murder_manslaughter_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = regression_data_no2states)
summary(model1)
par(mfrow = c(2,2))
plot(model1)

model2 = lm(robbery_rate ~ illicit_drugs + alcohol + marijuana + white + percent_high_school_graduate, data = regression_data_no2states)
summary(model2)
par(mfrow = c(2,2))
plot(model2)
```

The adjusted R-squared increased by removing outliers.

## Bootstrap

Use bootstrap to check the SE of two models

**Murder rate without Hawaii and District of Columbia**

```{r, echo=FALSE, warning=FALSE, message=FALSE}

set.seed(1)
# Try bootstrap with the model that has no outliers in murder rate
regression_data_no2states %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  filter(crime_type == "murder_manslaughter_rate") %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    linear_model = map(strap, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + percent_high_school_graduate, data = .x)),
    results = map(linear_model, broom::tidy)) %>% 
  select(results) %>%
  unnest() %>% 
  group_by(term) %>% 
  summarize(boot_se = sd(estimate))
```

**Murder rate with Hawaii and District of Columbia**

```{r, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(1)
# Try bootstrap with the model that has outliers in murder rate
regression_data_all %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  filter(crime_type == "murder_manslaughter_rate") %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    linear_model = map(strap, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + percent_high_school_graduate, data = .x)),
    results = map(linear_model, broom::tidy)) %>% 
  select(results) %>%
  unnest() %>% 
  group_by(term) %>% 
  summarize(boot_se = sd(estimate))
```

**Robbery rate without Hawaii and District of Columbia**

```{r, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(1)
# Try bootstrap with the model that has no outliers in robbery rate
regression_data_no2states %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  filter(crime_type == "robbery_rate") %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    linear_model = map(strap, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + percent_high_school_graduate, data = .x)),
    results = map(linear_model, broom::tidy)) %>% 
  select(results) %>%
  unnest() %>% 
  group_by(term) %>% 
  summarize(boot_se = sd(estimate))
```

**Robbery rate with Hawaii and District of Columbia**

```{r, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(1)
# Try bootstrap with the model that has outliers in robbery rate
regression_data_all %>%
  gather(key = crime_type, value = crime_rate, violent_crime_rate:motor_vehicle_theft_rate) %>%
  filter(crime_type == "robbery_rate") %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    linear_model = map(strap, ~lm(crime_rate ~ illicit_drugs + alcohol + marijuana + percent_high_school_graduate, data = .x)),
    results = map(linear_model, broom::tidy)) %>% 
  select(results) %>%
  unnest() %>% 
  group_by(term) %>% 
  summarize(boot_se = sd(estimate))
```

From the results, we found that all standard errors of coefficients in the model without the outliers were smaller than that of the model with outliers and we further confirmed that the model without the outliers (Hawaii, District of Columbia) was more accurate to predict the coefficients of the model.